Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2024-09-12 ms.qmd
Reformat for PLoS

Here, we use meta-analysis to establish the current state-of-knowledge of the extent to which aquatic and terrestrial ectotherms are capable of physiological plasticity. We then developed new effect sizes to quantify how variance in physiological rates changes with temperature to ask the following questions regarding acclimation-induced changes in trait means and variances: 1) Does variance in physiological rates change as temperatures rise? 2) Are temperature effects on means of physiological rates greater than changes in variance across aquatic and terrestrial ectotherms? 3) How do changes in trait mean and variance relate to different life-stages, traits, and habitats? 4) Are changes in mean and variance of physiological rates impacted by past climate history? 5) How are variances in physiological rates expected to change under climate change?

>>>>

Here, we use meta-analysis to establish the current state-of-knowledge of the extent to which aquatic and terrestrial ectotherms are capable of physiological plasticity. We developed new effect sizes ([@fig-concept]A)to quantify how variance in physiological rates change with temperature to ask the following questions regarding acclimation-induced changes in trait means and variances: 1) Does variance in physiological rates change as temperatures rise? 2) Are temperature effects on means of physiological rates greater than changes in variance across aquatic and terrestrial ectotherms? 3) How do changes in trait mean and variance relate to different life-stages, traits, and habitats? 4) Are changes in mean and variance of physiological rates impacted by past climate history? 5) How are variances in physiological rates expected to change under climate change? Importantly, unlike past analyses, our analysis corrects for changes in $Q_10$ that may vary across thermal performance curves across diverse taxa ([@fig-concept]B).
